---
title: "ProblemSet4_2"
author: 'Name: Zhen Qin, Uniqname: qinzhen, UMID: 48800866, Dept: Statistics'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Datasets are from https://github.com/jensqin/Stat506 .

## c.

```{r, message=FALSE, warning=FALSE}
setwd("E:/UM academy/stat 506/ProblemSet4")
library(sas7bdat)
library(lme4)

# load the data
rdclong = read.sas7bdat('rdclong.sas7bdat')

# Filter your data to contain only the 1000 Hz test for the right ear
test1kr = rdclong[rdclong$ear==1&rdclong$freq==1&rdclong$threshold<200,]
test = rdclong[rdclong$threshold<200,]
```

```{r}
# fit the model with interaction
fit_interaction=lm(threshold ~group*RIAGENDR, data=test1kr)
summary(fit_interaction)
```

The interaction is not significant because the p-value is 0.181>0.05.

```{r}
# fit the model to test age
fit_age=lm(threshold~RIDAGEYR+group+RIAGENDR+RIDAGEYR:group+RIDAGEYR:RIAGENDR,data=test1kr)
summary(fit_age)
```

After controlling for age group and gender, age is not important as a continuous variable because the p-value is 0.217>0.05.

```{r}
# fit the model to test group 
fit_group=lm(threshold~RIDAGEYR+group+RIDAGEYR:group,data=test1kr)
summary(fit_group)
```

The effect of age, as a continuous variable, is significantly different among the older and/or younger age groups because the p-value is less than 0.05.

## d.

```{r}
# fit the mixed model with interaction
mx_interaction=lmer(threshold ~group*RIAGENDR+(ear|freq),data = test)
summary(mx_interaction)
```

In fixed effects, the absolute value of t value of the interaction is greater than 2, so it is significant.

```{r}
# fit the mixed model to test age
mx_age=lmer(threshold~RIDAGEYR+group+RIAGENDR+RIDAGEYR:group+RIDAGEYR:RIAGENDR+(ear|freq),data=test)
summary(mx_age)
```

In fixed effects, the absolute value of t value of RIAGENDR is greater than 2, so it is significant.

```{r}
# fit the mixed model to test group
mx_group=lmer(threshold~RIDAGEYR+group+RIDAGEYR:group+(ear|freq),data=test)
summary(mx_group)
```

In fixed effects, the t-value is greater than 2 so the effect of age, as a continuous variable, is significantly different among the older and/or younger age groups.