---
title: "ProblemSet2_1"
author: 'Name: Zhen Qin, Uniqname: qinzhen, UMID: 48800866, Dept: Statistics'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####a.

First, I stored the data in `recs`.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
## Store data in recs
library('readr')
library('dplyr')
library('tidyr')
setwd('E:\\UM academy\\stat 506\\ProblemSet2')
recs=read_delim('./recs2009_public.csv', delim=',',col_names=TRUE)
```

Set variables. `Sweight` is sum of nweight grouped by `State` and `Roof`, `SSweight` is sum of nweight grouped by `State`. Generate the Rate matrix.

```{r}
## Calculate matrix including rate as Ratemat
select(recs, State=REPORTABLE_DOMAIN, Roof=ROOFTYPE, Weight = NWEIGHT) %>%
  group_by(State, Roof) %>%
  summarise(Sweight=sum(Weight)) %>%
  group_by(State) %>%
  mutate(SSweight=sum(Sweight),Rate=Sweight/SSweight) %>%
  select(Roof, State, Rate) %>%
  filter(Roof==2)->Ratemat
Ratemat=Ratemat[,-1]
```

Then showed `Ratemat` and calculate max value and min value of `Rate`.

```{r}
## Calculate max and min value
Statestr="Connecticut, Maine, New Hampshire, Rhode Island, Vermont
Massachusetts
New York
New Jersey
Pennsylvania
Illinois
Indiana, Ohio
Michigan
Wisconsin
Iowa, Minnesota, North Dakota, South Dakota
Kansas, Nebraska
Missouri
Virginia
Delaware, District of Columbia, Maryland, West Virginia
Georgia
North Carolina, South Carolina
Florida
Alabama, Kentucky, Mississippi
Tennessee
Arkansas, Louisiana, Oklahoma
Texas
Colorado
Idaho, Montana, Utah, Wyoming
Arizona
Nevada, New Mexico
California
Alaska, Hawaii, Oregon, Washington"
Statename=unlist(strsplit(Statestr,split = "\n"))
tmpmat=Ratemat
tmpmat$State=Statename
knitr::kable(tmpmat)
maxp=which.max(Ratemat$Rate)
minp=which.min(Ratemat$Rate)
plot(Ratemat$State,Ratemat$Rate,ylim = c(0,0.13))
axis(at=1:30,side = 1)
text(Ratemat$State[c(maxp,minp)],Ratemat$Rate[c(maxp,minp)],
     paste("(",Statename[c(maxp,minp)],",",round(Ratemat$Rate[c(maxp,minp)],4),")")
     ,pos=1,cex=0.7)
```

The max value is .1246495 and `State` is 16, the min value is .0144062 and `State` is 19. The answer 1 is North Carolina, South Carolina, the answer 2 is Tennessee.

####b.

Compute the proportion of each roof type for all houses constructed in each decade.

```{r}
## Compute the proportion
Roofstr="Not Applicable,Ceramic or Clay Tiles,Wood Shingles/Shakes,Metal,\
Slate or Synthetic Slate,Composition Shingles,Asphalt,Concrete Tiles,Other"
Roofname=unlist(strsplit(Roofstr,split = ","))
tmpstr=rep(Roofname[1],7)
for(i in 2:9){
  tmpstr=c(tmpstr,rep(Roofname[i],7))
}
Drecs=recs
Drecs$YEARMADERANGE[Drecs$YEARMADERANGE==8] = 7
select(Drecs, Roof=ROOFTYPE, Decade = YEARMADERANGE, Weight = NWEIGHT) %>%
  group_by(Roof, Decade) %>%
  summarise(Sweight=sum(Weight))->Smat
group_by(Smat,Decade) %>%
  mutate(SSweight=sum(Sweight),Rate=Sweight/SSweight) %>%
  select(Decade, Roof, Rate)->tmpmat
tmpmat$Decade=rep(c("before 1950","1950s",
                    "1960s","1970s","1980s","1990s","2000s"),9)
tmpmat$Roof=tmpstr
knitr::kable(tmpmat)
```

Calculate the max rate and generate a plot, and then use the string "Ceramic or Clay Tiles,Wood Shingles/Shakes,Metal,Slate or Synthetic Slate,Composition Shingles,Asphalt,Concrete Tiles,Other,Not Applicable" to get the names.

```{r, message=FALSE, warning=FALSE}
## Calculate the max rate and generate a plot
spread(Smat, Decade,Sweight) %>%
  transmute(Rate=(`2`+`3`+`4`+`5`+`6`)/`1`)->Ratemat
Roofname[which.max(Ratemat$Rate)]
plot(Ratemat$Roof,Ratemat$Rate,ylim = c(0,13),xlim = c(0,9))
axis(at=-2:8,side = 1)
text(Ratemat$Roof,Ratemat$Rate,paste("(",Roofname,",",round(Ratemat$Rate,4),")")
     ,pos=1,cex=0.7)
```

Use this subset by decades to calculate, Concrete Tiles roof saw the largest relative rise. 
