---
title: "STATS 551 - HW1"
author: "Zhen Qin"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 5.

Compute the total number of quizzes for 10000 times. The sample mean and sample variance of these numbers are very close to the theoretical values: mean = 13.5 and variance = 6.75. Thus the results in question 4 are verified.

```{r}
# set the number of trials
n = 10000

# create a function to generate results
quiz = function(){
  theta = runif(1,0,1)
  rbinom(1,1,theta)
}
total = rep(0,n)
for(i in 1:n){
  for(j in 1:27){
    total[i] = total[i] + quiz()
  }
}

# compute the sample mean and sample variance
mean(total)
var(total)
```


## Exploratory Data Analysis

The dataset is available [here](https://github.com/jensqin/stats551/blob/master/Crabs.dat.txt). There are six variables **crab**(the ID of the observed crab), **y**(stands for the number of crabs near the observation), **weight**, **width**, **color** and **spine**. 

### Description

First, I loaded the data using R and extracted 6 rows. It is clear that **y**, **weight** and **width** are numerical variables, while others are categorical variables.

```{r}
library(knitr)
library(ggplot2)
library(GGally)

# load the data and show 6 rows
crab = read.table("crabs.dat.txt", header = T)
kable(head(crab))
```

### Quantative Analysis

Second, I used quantative methods to show some properties. The median of **y** is less than the mean of **y**, which suggests **y** may have skewness. **weight** and **width** seem to have normality. 

```{r}
# data summary
crab = crab[,2:6]
crab$color = as.factor(crab$color)
crab$spine = as.factor(crab$spine)
kable(summary(crab))
```

Then I calculated the correlation matrix of numerical variables. It is worthwhile to note that the correlation of **weight** and **width** is big. Maybe they have colinearity.

```{r}
# correlation matrix of numerical variables
kable(cor(crab[1:3]))
```

### Graphic Analysis

Third, I used graphic methods to explore. 

The scatter plot matrix shows pairwise relationship of variables. This plot verifies the guess of quantative analysis.

```{r}
# scatter plot matrix
ggpairs(crab) + ggtitle("scatter plot")
```

The histogram of **y** shows there are many 0s in **y**, so I cannot use any model that assumes **y** has normality. This may be a potential difficult feature of the data.

```{r}
# histogram of y
ggplot(data = crab) + geom_histogram(aes(x = y), binwidth = 1) + ggtitle('Hist of y')
```

The regression line in the following plot has good fit, so these varibles have colinearity.

```{r}
# linearity between variables
ggplot(data = crab, aes(x = weight, y = width)) + geom_point(shape = 1) + geom_smooth(method = lm) + ggtitle('Linearity')
```

The boxplot shows **y** has considerable skewness when **color** is 2, 3 or 4.

```{r}
# boxplot by color
ggplot(data = crab, aes(x = color, y = y))+geom_boxplot()+ggtitle('Boxplot')
```

The freqency plot verifies this idea.

```{r}
# frequency plot
ggplot(data = crab, aes(x = y, y = ..density..))+geom_freqpoly(aes(colour = color),binwidth = 1)+ggtitle('Frequency')
```

The dot plot shows some categories have few cases, e.g. **color** = 1, **spine** = 3. Maybe they have relationship with **y** because all of them are not distributed evenly.

```{r}
# number of cases by color and spine
ggplot(data = crab, aes(x = color, y = spine)) + geom_count() + ggtitle('Count')
```

The scatter plot of MDS coordinates shows that points can be divided into 2 groups in a way.

```{r}
# MDS plot
mds = cmdscale(dist(crab,method = 'canberra'))
mds = as.data.frame(mds)
ggplot(data = mds, aes(x = V1, y = V2)) + geom_point() + ggtitle('MDS')
```

### Conclusion

Here are some interesting properties of the data:

* In **y** majority is equal to 0
* The two variables **weight** and **width** have colinearity. And they both have normality
* The categorical variables may be influential with respect to **y**

I will not use this dataset to do the final project. The main reason is that the data is relatively simple but I would like to do some challenging works. In addition, it is hard to get proper prior knowledge of crabs, so bayesian modeling may be not good for the data.