---
title: "ProblemSet3_1"
author: 'Name: Zhen Qin, Uniqname: qinzhen, UMID: 48800866, Dept: Statistics'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The data is from https://github.com/jensqin/Stat506.

#### a.

```{r, message=FALSE, warning=FALSE}
library('readr')
library(data.table)
setwd('E:\\UM academy\\stat 506\\ProblemSet3')

# load the data
recs=read_delim('./recs2009_public.csv', delim=',',col_names=TRUE)
recs=data.table(recs)[ROOFTYPE>0,]

# Store names of states
Statename= c("CT, ME, NH, RI, VT", "MA", "NY", "NJ", "PA", "IL",
             "IN, OH", "MI", "WI", "IA, MN, ND, SD", "KS, NE", "MO",
             "VA", "DE, DC, MD, WV", "GA", "NC, SC", "FL", "AL, KY, 
             MS","TN", "AR, LA, OK", "TX", "CO", "ID, MT, UT, WY",
             "AZ","NV, NM", "CA", "AK, HI, OR, WA" )

# Compute Rate matrix
recs$ROOFTYPE = as.factor(recs$ROOFTYPE)
levels(recs$ROOFTYPE) = c("Ceramic or Clay", "Wood Shingles", "Metal", "Slate", "Composition", 
                          "Asphalt", "Concrete", "Other")

Ratemat = recs[,.(State=REPORTABLE_DOMAIN,Roof=ROOFTYPE,Weight = NWEIGHT)
               ][, `:=`(Sweight={as.integer(Roof == "Wood Shingles")*Weight})
                 ][,.(Sweight = sum(Sweight), SSweight = sum(Weight)), by = State
                   ][, .(State, Rate = Sweight/SSweight)
                     ][order(State)]
Ratemat$State=Statename
Ratemat=Ratemat[order(Rate)]

# Show the table
knitr::kable(Ratemat, col.names = c('State(s)','Rate'))
```

1.

For each state(s) and each rooftype, compute sum of weight as Sweight.

2.

For each state, compute the sum of Sweight as SSweight, and compute the rate Sweight/SSweight.

3.

Draw state(s) and Statename, and sort the data by rate.

4.

Find which state(s) has the biggest rate, which has the smallest rate.

Answer: The answer 1 is North Carolina, South Carolina, the answer 2 is Tennessee.


#### b.

```{r}
# Compute the proportion of each roof type for all houses constructed in each decade
recs$YEARMADERANGE[recs$YEARMADERANGE==8]=7
recs$Decade = as.factor(recs$YEARMADERANGE)
levels(recs$Decade) = c("before1950", "1950s", "1960s", "1970s", "1980s", "1990s", "2000s")
sweight = recs[,.(Decade, ROOFTYPE, NWEIGHT)
                    ][, .(Sweight = sum(NWEIGHT)), by = Decade]
rate = recs[,.(Decade, ROOFTYPE, NWEIGHT)
                         ][, .(Weight = sum(NWEIGHT)), by = .(Decade, ROOFTYPE)]

# Set keys for computing
setkey(sweight, Decade)
setkey(rate, Decade)
rate = rate[sweight][, .(Prop = Weight/Sweight, Decade, ROOFTYPE)]
tmpmat = dcast(rate, ROOFTYPE ~Decade, value.var = "Prop")
tmpmat = tmpmat[,`:=`(rise=(`2000s`+`1990s`+`1980s`+`1970s`+`1960s`)/`before1950`)
          ][,.(Roof=ROOFTYPE, rise)]

# Show the table
knitr::kable(tmpmat)
```

1.

Combine the rows when yearmaderange is 7 or 8 as datasets in 2000s. Store a new attribute Decade.

2.

For each rooftype and decade, compute the sum of weight as sweight.

3.

For each decade, compute the sum of sweight, then compute the rate.

4.

Reshape the data into wide format, the formula is ROOFTYPE~Decade, value.var = "Prop", then compute relative rise in use between 1950 and 2000, draw the biggest rise.

Answer: Use this subset by decades to calculate, Concrete Tiles roof saw the largest relative rise.